---
title: "Week_1_Lab"
author: "JamesDale"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Source import
source("http://www.openintro.org/stat/data/arbuthnot.R")
source("http://www.openintro.org/stat/data/present.R")
```

## Exercise 1

```{r}
#scater plot number of female batisms over time
plot(x = arbuthnot$year, y = arbuthnot$girls)
```

```{r}
#line plot  number of female batisms over time
plot(x = arbuthnot$year, y = arbuthnot$girls, type = "l")
```

```{r echo=FALSE}
#Help link
#?plot
```

## Exercise 2
```{r}
#manual calculation
5218 + 4683
```

```{r}
#Adding boys and girls vectors
arbuthnot$boys + arbuthnot$girls
```

```{r}
#Plot showing the total number (boys+girls) of batisms per year
plot(arbuthnot$year, arbuthnot$boys + arbuthnot$girls, type = "l")
```

```{r}
#manual proportion of boys
5218 / 4683
```

```{r}
#Dividing vectors
arbuthnot$boys / arbuthnot$girls
```

```{r}
#Manual proportion of newborns that are boys
5218 / (5218 + 4683)
```

```{r}
#proportion of newborns that are boys
arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)
```

## Exercise 3

```{r}
#comparison operators
arbuthnot$boys > arbuthnot$girls
```

------

# Activities

## Activity 1
```{r}
#Years represented in data set
present$year
```

```{r}
#Name of variables
names(present)
```

```{r}
#Dimensions
dim(present)
```

## Exercise 2
### Comparisions:

#### The __difference__ in number of boys baptised in the arbuthnot data set and the present data set is : __`r format(abs(sum(arbuthnot$boys)-sum(present$boys)), scientific=FALSE)`__. The total in the arbuthnot data set is __`r if (sum(arbuthnot$boys) < sum(present$boys)) {"lower"} else { "higher"}`__ than in present.

#### The __difference__ in number of girls baptised in the arbuthnot data set and the present data set is : __`r format(abs(sum(arbuthnot$girls)-sum(present$girlss)), scientific=FALSE)`__. The total in the arbuthnot data set is __`r if (sum(arbuthnot$girls) < sum(present$girls)) {"lower"} else { "higher"}`__ than in present.

#### In the arbuthnot data set there is a greater number of  __`r if(sum(arbuthnot$boys)>sum(arbuthnot$girls)){"boys"}else{"girls"}`__.In the present data set there is a greater number of __`r if(sum(present $boys)>sum(present $girls)){"boys"}else{"girls"}`__

#### The data in the arbuthnot data set is spread over __`r max(arbuthnot$year)-min(arbuthnot$year)`__ years. The data in the present data set is spread over __`r max(present$year) - min(present$year)`__ years.


## Exercise 3
```{r}
plot(present$year, present$boys, xlab = "Years", ylab = "Number of Newborns", type = "l", col = "Blue", main = "Present")
par(new = TRUE)
plot(present$year, present$girls, type="l", xaxt = "n", yaxt = "n", ylab = "", xlab = "", col = "Red")
legend("topleft", legend=c("Boys", "Girls"),col=c("blue", "red"), lty = 1)
```

```{r}
plot(arbuthnot$year, arbuthnot$boys, xlab = "Years", ylab = "Number of Newborns", type = "l", col = "Blue", main = "Arbuthnot")
par(new = TRUE)
plot(arbuthnot$year, arbuthnot$girls, type="l", xaxt = "n", yaxt = "n", ylab = "", xlab = "", col = "Red")
legend("topleft", legend=c("Boys", "Girls"),col=c("blue", "red"), lty = 1)
```

#### Comparing the two graphs, we can see the ratio between girls and boys were about even in both the arbuthnot data and present data however, the ratio seems to deviate more in the arbuthnot data.

## Exercise 4
```{r}
present$births = present$boys + present$girls
```

#### The year with the highest number of births is __`r present[which.max(present$births),][1]`__
