---
title: "Week2"
author: "JamesDale"
date: "1/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("http://www.openintro.org/stat/data/cdc.R")
```

## Exercise 1
```{r print-head}
# Print head of df 
head(cdc)
```

### Summaries and Tables

```{r weight-summary}
# Summary of weight column 
summary(cdc$weight)
```

```{r r-calc}
#Using r as a calc
190 - 140
```

```{r built-in-functions}
# R built-in functions
mean(cdc$weight) 
var(cdc$weight)
median(cdc$weight)
```

```{r count-smokes}
# Count the responses in smoke100 col
table(cdc$smoke100)
```

```{r freq-distro}
#Relative frequency distribution
table(cdc$smoke100)/20000
```

```{r plt-smoke-bar}
#Plot smokeing column
barplot(table(cdc$smoke100))
```

```{r smoke-table-plot}
# Assign smoke variable and plot it
smoke <- table(cdc$smoke100)
barplot(smoke)
```


## Exercise 2
```{r ex2-table}
# Participants who have smoked across each gender
table(cdc$gender,cdc$smoke100)
```

```{r mosaic-plt-ex2}
# PLot table in mosaic 
mosaicplot(table(cdc$gender,cdc$smoke100))
```

## Exercise 3

```{r ex3-dim}
# Print dimensions of cdc data
dim(cdc)
```

```{r ex3-names}
# Print column names
names(cdc)
```

```{r ex3-indexing}
# SHow weights of first 10 responses
cdc[1:10,6]
```

```{r ex3-range}
# Range selection
1:10
```

```{r ex3-col-selc}
# Column indexing no range selection
cdc[,6]
```

```{r ex3-col-selc2}
# Column selection
cdc[,6]
```

```{r ex3-single-index}
# COlumn and single index selection
cdc$weight[567]
```

```{r ex3-col-and-range}
# COlumn and range selection
cdc$weight[1:10]
```

```{r ex3-conditional}
# Conditional statements 
cdc$gender == "m"
cdc$age > 30
```

```{r ex3-mdata-subset}
#Assign cdc where gender equals male to mdata
mdata <- subset(cdc, cdc$gender == "m")
# print head
head(mdata)
```

```{r ex3-over-and-below}
# assign variable to subset of cdc where gender is male and age over 30
m_and_over30 <- subset(cdc, gender == "m" & age > 30)
# assign variable to subset of cdc where gender is male or age over 30
m_or_over30 <- subset(cdc, gender == "m" | age > 30)
```

## Exercise 4

```{r ex4-plot-hiehgt}
# Plot height
boxplot(cdc$height)
```

```{r ex4-summary-height}
# Print summary of height
summary(cdc$height)
```

```{r ex4-boxplot}
# Compare height and gender with box plot
boxplot(cdc$height ~ cdc$gender)
```

```{r ex4-bmi-and-box}
# Assign bmi variable to weight divided by height squared multiplied by 703
bmi <- (cdc$weight / cdc$height^2) * 703
# PLot bmi verus general health
boxplot(bmi ~ cdc$genhlth)
```

## Exercise 5

```{r ex5-histo}
# Plot histogram of age
hist(cdc$age)
```

```{r ex5-bmi-hist}
# Plot histogram of BMI
hist(bmi)
# Plot histogram with small buckets/ticks
hist(bmi, breaks = 50)
```

## Activity 1

```{r act1-plot}
plot(cdc$weight ~ cdc$wtdesire)
```

#### As we can see from the scatter plot above, weight vs desired weight is fairly linear. This mean most respondents are at or near their disired weight. However there are a few interesting outliers. A few that are heavier than they desire and strangly a few that desire to be over 600.

## Activity 2

```{r act2-wdiff}
# Adding column wdiff (desired weight - current weight)
cdc$wdiff <- c(cdc$wtdesire - cdc$weight)
# Creating variable wdiff 
wdiff <- cdc$wdiff
```

## Activity 3

#### The data type of the column and variable wdiff is __`r typeof(wdiff)`__. If the result is 0 then the respondant is currently at their disired weight. If the result is negative the respondant wants to loose weight and if it is positive they would like to gain weight. 

## Activity 4
```{r wdiff-describe}
# Use this function to help describe wdiff
IQR = function(x) {
median_x <- median(x)
Q1 <- median(x[x<median_x])
Q3 <- median(x[x>median_x])
return(Q3-Q1)
}

# Create plots
par(mfrow=c(1,2))
hist(wdiff)
boxplot(wdiff)
```

#### The distrobution of wdiff is __uniform__ and __symmetric__.
* Mean: __`r mean(wdiff)`__
* Median: __`r median(wdiff)`__
* Standard Deviation: __`r sd(wdiff)`__
* IQR: __`r IQR(wdiff)`__

## Activity 5
```{r gender-comparision}
# Men and woman subsets
cdc_men <- subset(cdc, cdc$gender == "m")
cdc_women <- subset(cdc, cdc$gender == "f")

#Graphs
par(mfrow=c(1,3))
hist(cdc_women$wdiff)
hist(cdc_men$wdiff)
boxplot(wdiff ~ cdc$gender)

#Summaries
summary(cdc_men$wdiff)
summary(cdc_women$wdiff)
```

### Table
Equation | Men | Woman
-------- | ------- | -------
Mean | `r mean(cdc_men$wdiff)` | `r mean(cdc_women$wdiff)`
Median | `r median(cdc_men$wdiff)` | `r median(cdc_women$wdiff)`
Standard Deviation | `r sd(cdc_men$wdiff)` | `r sd(cdc_women$wdiff)`
IQR | `r IQR(cdc_men$wdiff)` | `r IQR(cdc_women$wdiff)`

#### Based on the graphs and summaries we can determine woman have a more skewed distrobution, a higher deviation, and want to loose more weight. 

## Activity 6
```{r one-deviation}
mean_weight <- mean(cdc$weight)
sd_weight <- sd(cdc$weight)
one_dev_below <- mean_weight - sd_weight
one_dev_above <- mean_weight + sd_weight
temp <- subset(cdc, weight > one_dev_below)
within_one_dev <- subset(temp, weight < one_dev_above)

#within_one_dev <- subset(temp, weight > one_dev_below && weight < one_dev_above ) NOT WORKING
```

#### The proportion of weights within one deviation of the mean is __`r (length(within_one_dev$weight)/length(cdc$weight)) * 100`__%.

